Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-14 doc/report/msc-report.tex
Procedure. sort of sucks, sort of not happy with this. please refactor at will.

TODO: Explain our working procedure regarding

\begin{itemize}
\item[-] Exploration
\item[-] Prototyping
\item[-] Real-world data
\item[-] Litterature studies
\item[-] Design phase
\end{itemize}

>>>>

This section describes how we structured our work over the course of the thesis
project. It describes how we, by working in an exploratory fashion, reached a
good understanding of the problem at hand, and how we successively inched closer
to a working solution by building early prototypes and iteratively improving
upon them.

The design phase, which we entered once our understanding of the problem scope
and its theoretical solution was satisfactory, and the following implementation
are described, followed by a summary of our literature studies.

\subsection*{Beginning}

% not so sure I like this section much..

We entered this project with a basic understanding of the problem Last.fm wanted
to solve but with little formal knowledge of the problem domain. We were tasked
with building a data store that would hopefully replace the store that was
presently in use.

We met the relevant people, were introduced to the current system and were
presented with literature in the area, namely OLAP Solutions and Whatever in
Data Warehousing.

\subsection*{Exploration by prototyping}

We explored the problem domain by asking questions about the environment, the
system we were to build and its requirements.

To explore the environment, for example, we asked what data sources there are
(web logs, scrobbling logs and radio streaming logs) and where that data is
stored (on HDFS, it turned out). We asked how much data the system was supposed
to handle (lots) and what the output format ought to be (any format, within
reason).

Most of the exploration was straight forward: many questions could be answered
from observations or by logical reasoning. Others were more open-ended, such as
deciding on a suitable output format. For the open-ended questions we would list
reasonable options and pitt them against each other, first on the drawing board
and eventually as prototypes.

% this is a little thin. The prototypes are small pieces of code that would
answer questions such as Is it at all possible to.. ? and What is the most
efficient way to..?

\subsection*{Design Phase}

After having become familiar with the problem and its theoretical solution, it
was time to design the software.

Our initial design was ambitious: it contained, amongst other things, a query
language and sub-components for query analyzation, to identity the most common
queries, and for query optimization. These are likely advantageous features in a
large data store but one can do without them.

In the end, we much simplified our initial design and ended up with a design
that was simpler and more to-the-point.

\subsection*{Implementation}

The implementation phase was rather small: we would simply stitch together our evolved prototypes and arrive at a solid ground to stand on: we had a system for running MapReduce jobs, storing data in HBase, etc.

Writing code on top of this that performed data cube manipulation was a kinch!

\subsection*{Literature studies}
% vet inte riktigt vad jag vill saga om det har.

Studying previous work in related fields brought OLAP systems to our attention.
It was to become very influential in our theoretical understanding of the
subject.

\subsection*{Summary}

This section has talked about the procedure we followed during our thesis work. It described our exploratory model and the prototyping we did, as well as the design phase that led up to the implementation.
